IML Term project paper

### 2022-12-11 paper.rmd
.

>>>>

### Support Vector Classifiers

The idea of the binary support vector classifier is based on the notion of finding within a $p$-dimensional problem a $(p-1)$-dimensional hyperplane $\lambda_0+\sum\lambda_i x_i=0$ to divide the set of explanatory variables. Now, for this hyperplane, we can divide our observations into two groups, those values for which $\lambda_0+\sum_{i=1}^m\lambda_j x_i>0$ and those values for which $\lambda_0+\sum\lambda_i x_i<0$. Thus, this can function as a model through which we can predict binary variables such as our variable of interest.

However, defining such hyperplanes can be slightly difficult as all values might not be separable by a hyperplane or the hyperplane that separates values can overfit to training data. Thus, we define a soft-margin classifier, which has the benefits of letting some values be misclassified or be closer to the hyperplane than a strict maximal margin classifier would define, which leads to a more robust model that fits most of the training dataset better.

In order to define this, we need to find the optimal hyperplane to separate the observations. This optimization problem is an optimization problem of $M$ in relation to coefficients $\lambda_i$ and slack variables $\epsilon$ within inequation $y_i(\lambda_0+\sum_{j=1}^m\lambda_jx_{ij})\ge M(1-\epsilon_i)$ with limitations $\sum_{j=1}^{m}\lambda_j^2=1$, $\epsilon_i\ge0$ and $\sum_{i=1}^n\epsilon_i\le C$ for tuning parameter $C$.

As in our optimization the data only depends on the inner products of the explanatory variables and the values of the variable of interest, we can use different generalizations of inner products to gain different kinds of models. Thus with different choices of kernels we can model different kinds of data.

Other benefits benefits of using support vector classifiers include that they have been proven to behave well in several different applications and that it has been shown that on average support vector classifiers perform better than logistic regression models in problems where the classes are well-separated. However, the model can vary quite notably based on the tuning parameter $C$ as this parameter affects the amount of observations that violate the margin and thus increases or decreases the amount of support vectors that influence the model.

For our model, we chose to run a 10-fold cross validation on a set of tuning parameters $C$ to assess which kernel performed best and chose to model the value based on that. In these preliminary assessments, we assessed the radial and linear kernels had the best performance of the ones available in the R package e1071. For these kernels, we ran loo-cross validation to assess the value of tuning parameter $C$ and calculated for said tuning parameter loo-cross-validated accuracy and perplexity. Eventually, the radial kernel had slightly better accuracy but the linear kernel had lower perplexity and as a result, we chose to use the linear kernel in our model
